### 2020-02-02 docs/materials/jsPsych/question/question4.Rmd

---
title: "question"
author: "Yoshihiko Kunisato"
output:
  html_document:
    mathjax: null
    highlight: null
    theme: null
    css: jspsych-6.1.0/css/jspsych.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
library(htmltools)
tagList(
tags$script(src='jspsych-6.1.0/jspsych.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-html-form.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-likert.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-multi-choice.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-multi-select.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-text.js')
)
```

```{js}
var multi_select = {
    type: 'survey-multi-select',
    questions: [
      {
        prompt: "好きな心理学を選んでください（複数回答可）", 
        options: ["知覚心理学", "学習心理学", "認知心理学", "発達心理学", "心理統計学","社会心理学","生理心理学","臨床心理学"], 
        horizontal: true,
        required: true,
        name: 'psych'
      }, 
      {
        prompt: "臨床心理学で好きな内容を選んでください（複数回答可）", 
        options: ["心理療法", "心理査定", "異常心理学"], 
        horizontal: false,
        required:　false,
        name: 'ClinicalPsych'
      }
    ], 
    button_label: '次へ',
    randomize_question_order: true
};

var timeline = [];
timeline.push(multi_select);

jsPsych.init({
    timeline: timeline,
    on_finish: function() {
     jsPsych.data.displayData();
    }
});
```