### 2020-02-02 materials/jsPsych/question/question1.Rmd

---
title: "question"
author: "Yoshihiko Kunisato"
output:
  html_document:
    mathjax: null
    highlight: null
    theme: null
    css: jspsych-6.1.0/css/jspsych.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
library(htmltools)
tagList(
tags$script(src='jspsych-6.1.0/jspsych.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-html-form.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-likert.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-multi-choice.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-multi-select.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-text.js')
)
```

```{js}
var form = {
  type: 'survey-html-form',
  preamble: '<p> 以下の質問に回答してください。</p>',
  html: '<p> 好きな食べ物: <input name="food_like" type="text" /></p>'+
  　　　'<p> 嫌いな食べ物: <input name="food_dislike" type="text" /></p>'+
        '<p> 好きなYoutuber: <input name="youtuber" type="text" /></p>',
  button_label: '次へ'
};

var timeline = [];
timeline.push(form);

jsPsych.init({
    timeline: timeline,
    on_finish: function() {
     jsPsych.data.displayData();
    }
});
```