### 2020-02-02 materials/jsPsych/question/question5.Rmd

---
title: "question"
author: "Yoshihiko Kunisato"
output:
  html_document:
    mathjax: null
    highlight: null
    theme: null
    css: jspsych-6.1.0/css/jspsych.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
library(htmltools)
tagList(
tags$script(src='jspsych-6.1.0/jspsych.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-html-form.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-likert.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-multi-choice.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-multi-select.js'),
tags$script(src='jspsych-6.1.0/plugins/jspsych-survey-text.js')
)
```

```{js}
var survey_text = {
  type: 'survey-text',
  questions: [
    {prompt: "氏名をご記入ください", name: 'name',placeholder: "姓 名"},
    {prompt: "jsPsychを学んだ感想を書いてください", rows: 10, columns: 80}, 
  ],
  button_label: '次へ',
};

var timeline = [];
timeline.push(survey_text);

jsPsych.init({
    timeline: timeline,
    on_finish: function() {
     jsPsych.data.displayData();
    }
});
```